---
title: "2. Multiple parameter configurations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2. Multiple parameter configurations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

One of the key contributions of the `flexurba` package is that it allows to *flexibly* reconstruct the Degree of Urbanisation classification. This means that the package enables users to adapt the standard specifications in the algorithm and construct alternative versions of the classification.

The following vignette demonstrates how `flexurba` can be used to construct alternative grid cell classifications by systematically varying different parameter settings.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Load the required data

As illustration, we will generate multiple variations of the grid cell classification for Belgium. First, we load the pre-processed population, built-up area and land grid in the data object `data_belgium`.

```{r setup}
library(flexurba)

# load the data
data_belgium <- load_grid_data_belgium()
```

### Create a list of parameter configurations

Next, we need to create a list with the different parameter configurations. Suppose we want to create a grid cell classification with each combination of the following parameters values:

-   An urban centre population density thresholds of 1500 and 2000:
    -   `UC_density_threshold = 1500`
    -   `UC_density_threshold = 2000`
-   An urban centre population size thresholds of 50 000 and 75 000:
    -   `UC_size_threshold = 50000`
    -   `UC_size_threshold = 75000`
-   Urban centres constructed with gap filling and edge smoothing, and without gap filling and edge smoothing
    -   `UC_gap_fill = TRUE` and `UC_edge_smooth = TRUE`
    -   `UC_gap_fill = FALSE` and `UC_edge_smooth = FALSE`

This can be established with the following code. First, the standard DEGURBA parameter settings are loaded in the data object `standard_param`.

```{r}
standard_param <- get_grid_parameters()
```

Next, we create variables with the different parameter values, as defined above:

```{r}
UC_density_variations <- c(1500, 2000)
UC_size_variations <- c(50000, 75000)
UC_gapsmooth_variations <- c(TRUE, FALSE)
```

Finally, a list is created with the final parameter configurations by inserting the alternative parameter values in the object `standard_param` with the standard parameter settings:

```{r}
# create an empty list to save the different parameter configurations
parameter_variations <- list()

# iterate over each combination of parameter values
# and save the configuration in the list
i <- 1

for (density in UC_density_variations) {
  for (size in UC_size_variations) {
    for (gapsmooth in UC_gapsmooth_variations) {
      # get standard parameters
      parameters <- standard_param

      # adapt the parameters values
      parameters$UC_density_threshold <- density
      parameters$UC_size_threshold <- size
      parameters$UC_gap_fill <- gapsmooth
      parameters$UC_smooth_edge <- gapsmooth

      # save in the list
      parameter_variations[[i]] <- parameters

      i <- i + 1
    }
  }
}
```

### Execute the different parameter configurations

Finally, the grid classification with each of these parameter configurations is executed and plotted as follows:

```{r}
# create a list to save the classifications
classifications <- list()
plots <- list()

# create the grid classification and plot the result
for (i in seq_along(parameter_variations)) {
  classifications[[i]] <- classify_grid(
    data = data_belgium,
    parameters = parameter_variations[[i]]
  )

  plots[[i]] <- plot_grid(classifications[[i]],
    title = paste0(
      "Grid classification (", parameter_variations[[i]]$UC_density_threshold, ", ",
      parameter_variations[[i]]$UC_size_threshold, ", ",
      parameter_variations[[i]]$UC_gap_fill, ")"
    )
  )
}
plots
```
