---
title: "1. Degree of Urbanisation: Level 2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Degree of Urbanisation: Level 2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The Degree of Urbanisation methodology includes two hierarchical levels of classification. **Level 1 of DEGURBA** distinguishes (1) cities, (2) towns and semi-dense areas, and (3) rural areas. In **Level 2 of DEGURBA**, the classes in Level 1 are further subdivided into a total of 7 different classes. Cities are not subdivided. Towns and semi-dense areas are divided in (1) dense towns, (2) semi-dense towns and (3) suburban or peri-urban areas. Rural areas are divided in (1) villages, (2) dispersed rural areas and (3) mostly uninhabited areas.

Level 2 is developed to capture the full rural-urban continuum, and to provide insight into the spatial structure of a certain region (see [methodological manual](https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Applying_the_degree_of_urbanisation_manual) for more information).

This vignette describes how the `flexurba` package can be used to establish both the grid cell classification and the spatial units classification according to Level 2 of the Degree of Urbanisation.

## Grid classification of Level 2

The grid classification in (1) urban centres, (2) urban clusters and (3) rural grid cells in Level 1 of DEGURBA is further subdivided as follows:

-   **Urban centres** are not subdivided.
-   **Urban clusters** are subdivided in (1) dense urban clusters, (2) semi-dense urban clusters and (3) suburban and peri-urban clusters.
-   **Rural grid cells** are subdivided in (1) rural clusters, (2) low density rural cells and (3) very low density rural cells.

For the exact classification rules, see the documentation of the function `DoU_classify_grid()` or the official [DEGURBA methodological manual](https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Applying_the_degree_of_urbanisation_manual).

The grid classification of Level 2 can be constructed with the following code. First, we need to load the required data. In pre-processed population, built-up area and land grid are loaded in the object `data_belgium`:

```{r}
library(flexurba)

# load the data for Belgium
data_belgium <- DoU_load_grid_data_belgium()
```

Afterwards, the classification with the standard parameter settings can be constructed as follows.

```{r}
#| fig.alt: >
#|   Grid classification with the standard parameters
# classification with the standard parameters
classification1 <- DoU_classify_grid(
  data = data_belgium, 
  level1 = FALSE
)

# plot the classification
DoU_plot_grid(
  classification1, 
  level1 = FALSE
)
```

However, the `flexurba` package has more functionalities. The function `DoU_classify_grid()` allows the user to adapt various parameters in the classification algorithm, including the the minimum population thresholds or the contiguity rules.

As an example, the code below changes four different parameters in the Level 2 classification algorithm:

-   `UC_density_threshold` and `DUC_density_threshold`: we set the minimum population density threshold of urban centres (`UC`) and dense urban clusters (`DUC`) to 1000 inhabitants per km² (instead of the standard value of 1500 inhabitants per km²).

-   `DUC_contiguity_rule`: we change the contiguity rule for dense urban clusters (`DUC`) from the standard queen contiguity (`8`) to rook contiguity (`4`).

-   `RC_size_threshold`: the minimum size threshold for rural clusters (`RC`) is set to 400 inhabitants (instead of the standard threshold of 500 inhabitants).

```{r}
#| fig.alt: >
#|   Grid classification with the custom parameters
# specify custom parameters
classification2 <- flexurba::DoU_classify_grid(
  data = data_belgium,
  level1 = FALSE,
  parameters = list(
    UC_density_threshold = 1000,
    DUC_density_threshold = 1000,
    DUC_contiguity_rule = 4,
    RC_size_threshold = 400
  )
)

# plot the classification
DoU_plot_grid(
  classification2, 
  level1 = FALSE
)
```

The code demonstrates just four different examples of parameters that can be adjusted in the algorithm. However, the function `DoU_classify_grid()` incorporates many other possible parameters, including the built-up threshold value and the gap filling and smoothing rules. An overview of all possible parameters and their standard values can be found in the documentation of the function `DoU_classify_grid()` in the section 'Custom specifications'.

## Spatial units classification of Level 2

The Level 1 spatial units classification in (1) cities, (2) towns and semi-dense areas, and (3) rural areas is further subdivided as follows:

-   **Cities** are not subdivided
-   **Towns and semi-dense areas** are subdivided in (1) dense towns, (2) semi-dense towns and (3) suburbs or peri-urban areas.
-   **Rural areas** are subdivided in (1) villages, (2) dispersed rural areas and (3) mostly uninhabited areas.

For the exact classification rules, see the documentation of the function `DoU_classify_units()` or the official [DEGURBA methodological manual](https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Applying_the_degree_of_urbanisation_manual). The code below illustrates how to construct the units classification of Level 2.

Spatial units classification at municipal level can be classified as follows.

```{r}
#| fig.alt: >
#|   Units classification at municipal level
# pre-process the data
data_units <- DoU_preprocess_units(
  units = flexurba::units_belgium,
  classification = classification1,
  pop = data_belgium$pop
)

# get the units classification
units_classification <- DoU_classify_units(
  data_units, 
  level1 = FALSE
)

# plot the classification
plot_units(
  units = data_units$units, 
  classification = units_classification,
  level1 = FALSE
)
```

Furthermore, the function `DoU_classify_units()` allows the user to classify units on a different level of aggregation. For example, by aggregating the Belgian municipalities to districts (*Dutch: arrondissementen*) based on a common attribute field (in this dataset: `GID_3`).

```{r}
#| fig.alt: >
#|   Units classification at district level
# pre-process the data, dissolve by level 3 (= Belgian districts)
data_units2 <- DoU_preprocess_units(
  units = flexurba::units_belgium,
  classification = classification1,
  pop = data_belgium$pop,
  dissolve_units_by = "GID_3"
)

# get the units classification
units_classification2 <- DoU_classify_units(
  data_units2, 
  level1 = FALSE, 
  id = "GID_3"
)

# plot the classifications
plot_units(
  units = data_units2$units, 
  classification = units_classification2, 
  level1 = FALSE
)
```
